---
title: "New Function Section"
author: "Thomas Nguyen"
date: "Sys.Date()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r include = FALSE}
library(tidyverse)
```

```{r}
tb_long <- readRDS("tb_long.rds")
```
Last week, we were introduced with R function and its component. This week, we studied about ggplot2 and wrangling verb. Now, let's combine them together.

 
## Why should you use function?
Function is a powerful way to automate your tasks. You will become more productive using function.  

1. Function enable easy reuse within a project, helping you not to repeat yourself. Use it whenever you copy and paste a block of code more than twice.  
2. Function eliminates the chance of making incidental mistakes when you copy and paste (i.e. updating a variable name in one place, but not in another).   
3. Function can be given evocative name that makes your code easier to understand.  
4. Function will help you to effortlessly update your code as requirements change.  


```{r example, eval = FALSE}
# Do you see any problems in this code chunk?
# Too much repetition?
tb_long %>% filter(country == "Australia") %>%
  ggplot(aes(x = year, y = count))+
  geom_col()


tb_long %>% filter(country == "Brazil") %>%
  ggplot(aes(x = year, y = count))+
  geom_col()


tb_long %>% filter(country == "China") %>%
  ggplot(aes(x = year, y = count))+
  geom_col()

```


## How to write function:

```{r}
function(inputs){
  
  # code
  # more code
  
  return(outputs)
  }
```

1. Pick a short but informative name, preferably a verb.  
2. List inputs, or arguments, inside the round brackets inside function. If we had more the call would look like function(x, y, z).    
3. Place the code you have developed in body of the function, a {} block that immediately follows function(...).    

Note that, is is easier to start with working code and turn it into a function; it’s harder to create a function and then try to make it work.


Example:

```{r}
# Function to make a bar chart to see the number of cases of tuberculosis across years for a specific country.
# Identify inputs: data and a country name.
# First argument (x) is data
# Second argument (y) is the country we want to draw the graph for (Must be inside double quotation mark)

# Simple function
plot_a_country <- function(x, y){
  # Filter the country
  plot_dat <- x %>% filter(country == y)
  # Draw the plot
  plot <- plot_dat %>% 
    ggplot(aes(x = year, y = count))+
    geom_col()
  
  return(plot)
  }

# Checking the function
plot_a_country(tb_long, "China")


# Make our function easier to read
# And change the color
plot_a_country <- function(data, country_name){
  # Filter the country
  plot_dat <- data %>% filter(country == country_name)
  # Draw the plot
  plot <- plot_dat %>% 
    ggplot(aes(x = year, y = count))+
    geom_col(fill = "steelblue")
  
  return(plot)

}
# Checking the function
# The function provides a blank plot since "USA" is not a valid country name
# Thefore no data is filtered and provided to plot. 
plot_a_country(tb_long, "USA")

# The function works as "United States of America" is a valid country name.
plot_a_country(tb_long, "United States of America") 

# You can further modify the plot after using the function
plot_a_country(tb_long, "United States of America") +
  theme_bw()+
  ggtitle("USA tuberculosis by year")



# Note: I made the function to return the plot object so it is more pipeable
# (I can add plot extras & further modify my plot later). 
# For example, “+ theme()”or “+scale_colour_brewer()” can be added further down the pipe.

```


## About curly curly => {{}} 

It was hard to write function with tidyverse wrangling verbs. R will throw all type of error at you. You couldn't easily use select(), summarise(), group_by() inside a function. Curly curly helps with that. It work *within* the function.

When you create a function using select(), summarise(), group_by(), wrap the function arguments containing data frame variables with {{.

Example:

```{r eval = FALSE}
# This code will run nicely

select_run <- function(data, var1, var2){
   data %>% select({{var1}}, {{var2}}) 
 }
select_run(tb_long, type, year)

# This will happen if you remove the curly curly 

 select_notrun <- function(data, var1, var2){
   data %>% select(ar1, var2) 
 }
 select_notrun(tb_long, type, year)
```

```{r}
# You can not run this function if you remove the curly curly

# Function to find the total cases by country since first recorded
 
total_by <- function(data, var, by) {
  data %>%
    group_by({{by}}) %>%
    summarise(total_cases = sum({{var}}, na.rm = TRUE)) %>%
    arrange(desc(total_cases))
}

total_by(tb_long,count, country)
```

