---
title: "Function_Tut"
author: "Thomas Nguyen"
date: "Sys.Date()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r include = FALSE}
library(tidyverse)
```

```{r data}
tb_long <- readRDS("tb_long.rds")
```

### How to write function:

```{r}
myfun <- function(inputs){outputs}

myfun <- function(inputs){
  # Codes
  # Lots of code
  
  return(output)
}
```



```{r}

# A few tips
# 1. Pick a short but informative name, preferably a verb.  
# 2. List inputs, or arguments, inside the round brackets inside function. If we had more the call would look like function(x, y, z).    
# 3. Place the code you have developed in body of the function, a {} block that immediately follows function(...). Remember to document them.

```

Example:


```{r}
# Function to make a bar chart to see the number of cases of tuberculosis across years for a specific country.
# First argument (x) is data
# Second argument (y) is the country we want to draw the graph for (Must be inside double quotation mark)
# X axis demonstrates year
# Y axis demonstrates count

tb_long %>%
  filter(country == "Australia") %>%
  ggplot(aes(x = year, y = count))+
  geom_col(fill = "blue")

tb_long %>%
  filter(country == "China") %>%
  ggplot(aes(x = year, y = count))+
  geom_col(fill = "blue")


plot_a_country <- function(data, country_name){
  # Filter country
  plot_dat <- data %>% filter(country == country_name)
  # Draw the plot
  plot <- plot_dat %>%   
    ggplot(aes(x = year, y = count))+
    geom_col(fill = "red")
  
  return(plot)
}

plot_a_country(tb_long, "China")
plot_a_country(tb_long, "Australia")
plot_a_country(tb_long, "Belarus")


```


## About curly curly => {{ }} 

It was hard to write function with tidyverse. R will throw all type of error at you. You couldn't easily use select(), summarise(), group_by() inside a function. Curly curly helps with that. It work *within* the function.

When you create a function using select(), summarise(), group_by(), wrap the function arguments containing data frame variables with {{.

Example:

```{r eval = FALSE}
# This code will run nicely

select_run <- function(data, var1, var2){
   data %>% select({{var1}}, {{var2}}) 
 }

select_run(tb_long, type, year)

# This will happen if you remove the curly curly 

 select_notrun <- function(data, var1, var2){
   data %>% select(var1, var2) 
 }
 select_notrun(tb_long, type, year)
```

```{r}
# You can not run this function if you remove the curly curly

# Function to find the total cases by country since first recorded
 
total_by <- function(data, var, by) {
  data %>%
    group_by({{by}}) %>%
    summarise(total_cases = sum({{var}}, na.rm = TRUE)) %>%
    arrange(desc(total_cases))
}

total_by(tb_long,count, country)
```

